\documentclass[10pt,landscape]{article}
\usepackage{multicol}
\usepackage{calc}
\usepackage{ifthen}
\usepackage[landscape]{geometry}
\usepackage{hyperref}
\usepackage{cite}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{float}

% To Do:
% \listoffigures \listoftables
% \setcounter{secnumdepth}{0}


% This sets page margins to .5 inch if using letter paper, and to 1cm
% if using A4 paper. (This probably isn't strictly necessary.)
% If using another size paper, use default 1cm margins.
\ifthenelse{\lengthtest { \paperwidth = 11in}}
	{ \geometry{top=.5in,left=.5in,right=.5in,bottom=.5in} }
	{\ifthenelse{ \lengthtest{ \paperwidth = 297mm}}
		{\geometry{top=1cm,left=1cm,right=1cm,bottom=1cm} }
		{\geometry{top=1cm,left=1cm,right=1cm,bottom=1cm} }
	}

% Turn off header and footer
\pagestyle{empty}
 

% Redefine section commands to use less space
\makeatletter
\renewcommand{\section}{\@startsection{section}{1}{0mm}%
                                {-1ex plus -.5ex minus -.2ex}%
                                {0.5ex plus .2ex}%x
                                {\normalfont\large\bfseries}}
\renewcommand{\subsection}{\@startsection{subsection}{2}{0mm}%
                                {-1explus -.5ex minus -.2ex}%
                                {0.5ex plus .2ex}%
                                {\normalfont\normalsize\bfseries}}
\renewcommand{\subsubsection}{\@startsection{subsubsection}{3}{0mm}%
                                {-1ex plus -.5ex minus -.2ex}%
                                {1ex plus .2ex}%
                                {\normalfont\small\bfseries}}
\makeatother

% Define BibTeX command
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

% Don't print section numbers
\setcounter{secnumdepth}{0}


\setlength{\parindent}{0pt}
\setlength{\parskip}{0pt plus 0.5ex}


% -----------------------------------------------------------------------

\begin{document}

\raggedright
\footnotesize
\begin{multicols}{3}


% multicol parameters
% These lengths are set only within the two main columns
%\setlength{\columnseprule}{0.25pt}
\setlength{\premulticols}{1pt}
\setlength{\postmulticols}{1pt}
\setlength{\multicolsep}{1pt}
\setlength{\columnsep}{2pt}

\begin{center}
     \Large{\textbf{JAVA Cheat Sheet 4\\Data Structures and Advanced Concepts}} \\
\end{center}

\section{Data Structures}
\subsection{Arrays}

\begin{itemize}
\item Objects that store multiple variables of the same type.
\item Can hold either primitives or object references.
\item Will always be an object on the heap.
\end{itemize}
\subsubsection{Declaration}
\begin{lstlisting}
// common
int[] key; 
// better avoid due to readability
int key [];
\end{lstlisting}
\subsubsection{Construction}
Creates the object on the heap.\\

\begin{lstlisting}
// The size has to be specified
int[] myArray = new int[3];
\end{lstlisting}

For multidimensional arrays omitting the second dimension's size is possible:\\

\begin{lstlisting}
int[][] myArray = new int[3][];
\end{lstlisting}
\subsubsection{Initializing}
Arrays can be initialized by assigning values to their elements:\\
\begin{lstlisting}
int[] myArray = int[3];
myArray[0] = 5;
\end{lstlisting}
The index always starts with 0 and the highest index equals the array's length -1.\\
Declaration, construction and initialization is also possible in one single step:\\
\begin{lstlisting} 
int[] dots = {1,3,5}
\end{lstlisting}

\subsubsection{Anonymous Array Creation}
It is possible to create an anonymous array, for example to pass it as a functions parameter.\\
\begin{lstlisting}
functionCall(new int[]{1,2,3});
\end{lstlisting}

\subsection{Enumerations}
The JAVA type for enumerations is called enum.\\
The restrict the values for a variable to a predefined set of values.\\
\begin{lstlisting}
enum CoffeeSize{
  BIG, HUGE, OVERWHELMING
};
\end{lstlisting}
The ';' at the end is optional.\\
An enum can be seen as a special kind of class. The values are constants (public static final) instances of this class. One can define constructors and other methods:
\begin{lstlisting}
enum CoffeeSize{
  BIG(8), HUGE(10), OVERWHELMING(16);
  CoffeeSize(int ounces){
    this.ounces = ounces;
  }
  private int ounces;
  private int getOunces(){
    return this.ounces;
  }
}
\end{lstlisting} 
Enums also can have a constant specific class body which overrides a method:\\
\begin{lstlisting}
  ...
  OVERWHELMING(16){
    public String getLidCode(){
      return "B";
    }
  };
  ...
\end{lstlisting}

The enum-class implements equals and hashCode, therefore it is possible to use enums as keys for maps.

\subsection{Lists, Trees, Collections}
In JAVA we distinguish between \textbf{collections}, a \textbf{Collection} and the utility class \textbf{Collections}.\\
There are many classes considered to fulfil the concept of a collection. Those are collections. Then there is the class Collection, which is the interface for a subset of these collections. And then there is a utility class providing static methods for collections.\\
The single collections cannot take primitives as elements. In this case one have to use the wrapper types. Java implicitly uses \textbf{auto-boxing}, so it is possible to give primitives as arguments instead of their wrapper type representations.
\subsubsection{Overview over the collections}
\begin{itemize}
	\item Collection
	\begin{itemize}
		\item Set
		\begin{itemize}
			\item HashSet
			\begin{itemize}
				\item LinkedHashSet
			\end{itemize}
			\item SortedSet
		\end{itemize}
		\item List
		\begin{itemize}
			\item ArrayList
			\item Vector
			\item LinkedList
		\end{itemize}
		\item Queue
		\begin{itemize}
			\item LinkedList
			\item PriorityQueue
		\end{itemize}
	\end{itemize}
	\item Map
	\begin{itemize}
		\item HashTable
		\item HashMap
		\begin{itemize}
			\item LinkedHashMap
		\end{itemize}
		\item SortedMap
		\begin{itemize}
			\item Navigable Map
			\begin{itemize}
				\item TreeMap
			\end{itemize}
		\end{itemize}
	\end{itemize}
\end{itemize}

\subsubsection{Lists}
There are three implementations of the interface List. They all provide the common methods of Collection and additionally provide methods related to the index of an element.\\
\textbf{ArrayList}\\
An ArrayList provides fast access and fast iterations over its elements.\\
But every time its allocated memory is exceeded by inserting new elements a new list has to be created and the whole old list has to copied into the new list.\\
\begin{lstlisting}
List<String> list = new ArrayList<String>();
list.add("hello");
String string = list.get(0);
\end{lstlisting}
\textbf{Vector}\\
Leftover from earlier JAVA versions. Mainly the same as an ArrayList, but uses synchronized methods for the use in multi-threading, which makes it slower.\\
\begin{lstlisting}
List<String> list = new Vector<String>();
list.add("hello");
String string = list.get(0);
\end{lstlisting}
\textbf{LinkedList}\\
A double-linked list. Provides methods for adding from the beginning or the end, which makes it a good choice for queues or stacks. Insertion and Deletion is faster than for the other two lists, but iteration and access may be slower.
\begin{lstlisting}
List<String> list = new LinkedList<String>();
list.add("hello");
String string = list.get(0);
\end{lstlisting}
\subsubsection{Trees}
There are only two structures using trees:
\begin{itemize}
\item TreeSet
\item TreeMap
\end{itemize}
Both provide an ascending order for the elements based on the elements' natural order.\\
The methods \textbf{lower}($<$), \textbf{floor}($\leq$) and \textbf{higher}($>$), \textbf{ceiling}($\geq$) can be used to navigate or search within tree structures.\\

\subsubsection{Collections}
This class contains utility methods for collections, for example to search elements, sort collections, reverse the order of a list and so on.

\section{Advanced Concepts}
\subsection{Iterators}
An \textbf{iterator} is an object that lets one loop over a collection step by step.\\
\begin{lstlisting}
Iterator<Dog> iterator = d.iterator();
while(iterator.hasNext()){
  Dog dog = iterator.next();
  System.out.println(dog.name);
}
\end{lstlisting}

\subsection{Recursion}

\subsection{Templates}

\subsection{Generics}

\subsection{Reflection}

\subsection{Annotation}

\end{multicols}
\end{document}
